---
title: "Correlation_Similarity"
author: "Thomas Naake"
date: "2023-01-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the libraries

```{r}
library(tidyr)
library(dplyr)
library(ComplexHeatmap)
library(ggplot2)
```

## Load the dataset

Load the dataset.

```{r }
df_all <- readr::read_delim("~/GitHub/SimilarityHackathon/data/parameter_tests_NEG_full.csv")

## calculate from distance to similarity for Euclidean/Avdist
df_all[, grepl("neuclidean", colnames(df_all))] <- 1 - df_all[, grepl("neuclidean", colnames(df_all))]
df_all[, grepl("navdis", colnames(df_all))] <- 1 - df_all[, grepl("navdis", colnames(df_all))]

## remove the rows where we don't have tanimoto_scores
df_all <- df_all |>
    filter(!is.na(tanimoto_scores))
```

## All the dataset

```{r}
## remove the pccid and the matching column
df <- df_all |>
    select(-c(pccid, matching))

hist(df$tanimoto_scores, main="Tanimoto coefficients (all)")
```


## Correlation between similarity coefficients and Tanimoto

`cols` specifies the columns that contain the Spectral similarity scores, 
not the Tanimoto coefficients.

### Calculate correlation

```{r}
## for Pearson correlation
.cols_remove <- c("id_query", "ref_id") 
corr_pearson <- df |>
    select(-all_of(.cols_remove)) |>
    as.matrix() |>
    cor(method = "pearson", use = "complete.obs")

## for Spearman correlation
corr_spearman <- df |> 
    select(-all_of(.cols_remove)) |>
    as.matrix() |>
    cor(method = "spearman", use = "complete.obs")

## extract the information on the Tanimoto coefficient and delete the 
## respective columns and rows

.col <- "tanimoto_scores" ############################################
corr_pearson_tanimoto <- corr_pearson[, .col]
corr_spearman_tanimoto <- corr_spearman[, .col]

.ind_remove <- !colnames(corr_pearson) %in% .col
corr_pearson <- corr_pearson[.ind_remove, .ind_remove]
corr_spearman <- corr_spearman[.ind_remove, .ind_remove]
```


Add annotations for rows.

```{r}
.rn <- rownames(corr_pearson)
algorithm <- character(length = nrow(corr_pearson))
tolerance <- numeric(length = nrow(corr_pearson))
mz_power <- numeric(length = nrow(corr_pearson))
intensity_power <- numeric(length = nrow(corr_pearson))
allowed_missing_percentage <- rep(NA, length = nrow(corr_pearson))

## algorithm
algorithm <- ifelse(grepl("nspecangle", .rn), "Spectral_angle", algorithm)
algorithm <- ifelse(grepl("cosine", .rn), "Cosine", algorithm)
algorithm <- ifelse(grepl("navdis", .rn), "Absolute_distance", algorithm)
algorithm <- ifelse(grepl("neuclidean", .rn), "Euclidean", algorithm)
algorithm <- ifelse(grepl("Neutralloss", .rn), "Neutral_loss", algorithm)
algorithm <- ifelse(grepl("Spec2Vec", .rn), "Spec2Vec", algorithm)
algorithm <- algorithm |>
    as.character()

## tolerance
tolerance <- ifelse(grepl("nspecangle", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 8)), tolerance)
tolerance <- ifelse(grepl("cosine", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 8)), tolerance)
tolerance <- ifelse(grepl("navdis", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 8)), tolerance)
tolerance <- ifelse(grepl("neuclidean", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 8)), tolerance)
tolerance <- ifelse(grepl("Neutralloss", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 2)), tolerance)
tolerance <- ifelse(grepl("Spec2Vec", .rn), NA, tolerance)
tolerance <- tolerance |>
    as.numeric()

## mz_power
mz_power <- ifelse(grepl("nspecangle", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), mz_power)
mz_power <- ifelse(grepl("cosine", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), mz_power)
mz_power <- ifelse(grepl("navdis", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), mz_power)
mz_power <- ifelse(grepl("neuclidean", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), mz_power)
mz_power <- ifelse(grepl("Neutralloss", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 3)), mz_power)
mz_power <- ifelse(grepl("Spec2Vec", .rn), NA, mz_power)
mz_power <- mz_power |>
    as.numeric()

## intensity_power
intensity_power <- ifelse(grepl("nspecangle", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 6)), intensity_power)
intensity_power <- ifelse(grepl("cosine", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 6)), intensity_power)
intensity_power <- ifelse(grepl("navdis", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 6)), intensity_power)
intensity_power <- ifelse(grepl("neuclidean", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 6)), intensity_power)
intensity_power <- ifelse(grepl("Neutralloss", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), intensity_power)
intensity_power <- ifelse(grepl("Spec2Vec", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 2)), intensity_power)
intensity_power <- intensity_power |>
    as.numeric()


## allowed_missing_percentage
## Spec2Vec


column_ha <- HeatmapAnnotation(algorithm = algorithm)
row_ha <- rowAnnotation( 
    tolerance = as.factor(tolerance), 
    mz_power = as.factor(mz_power), 
    intensity_power = as.factor(intensity_power))

## Pearson coefficients
Heatmap(mat = corr_pearson, top_annotation = column_ha,
    right_annotation = row_ha, show_column_names = FALSE, show_row_names = FALSE)

## Spearman coefficients
Heatmap(mat = corr_spearman, top_annotation = column_ha,
    right_annotation = row_ha, show_column_names = FALSE, show_row_names = FALSE)
```

### Scatterplots



Using `ggplot2`

```{r}
##sample_prop <- sample(1:nrow(df), ceiling(nrow(df) * 0.2), 
##    replace = FALSE)


df |>##[sample_prop, ] |>
    ggplot(aes_string(x = colnames(df)[3], y = colnames(df)[4])) +
    geom_point(cex = 0.2, alpha = 0.2) +
    geom_smooth(method = "lm") +
    theme_classic()

df |>##[sample_prop, ] |>
    ggplot(aes_string(x = colnames(df)[3], y = colnames(df)[99])) +
    geom_point(cex = 0.2, alpha = 0.2) +
    geom_smooth(method = "lm") +
    theme_classic()
```


### Tanimoto against spectral similarities

The last column (`ncol(df)`) contains the Tanimoto score.

```{r}
df_l <- pivot_longer(df, cols = 4:ncol(df))
```

Do the actual plotting. Only show the data for the top 5 scores

```{r}
corr_spearman_tanimoto_sort <- sort(corr_spearman_tanimoto, decreasing = TRUE)

corr_spearman_tanimoto_sort[1:10]


## remove the .col (Tanimoto column) with correlation of 1
corr_spearman_tanimoto_sort <- corr_spearman_tanimoto_sort[!names(corr_spearman_tanimoto_sort) %in% .col]

## create histograms per algorithm that show the rank of the correlation
df_ranks <- data.frame(
    corr = sort(corr_spearman_tanimoto_sort, decreasing = T),
    ranks = order(corr_spearman_tanimoto_sort, decreasing=T),
    names = names(corr_spearman_tanimoto_sort[order(corr_spearman_tanimoto_sort, decreasing=T)]),
    tolerance = NA,
    algorithm = "")

## fill the tolerance
df_ranks$tolerance <- ifelse(grepl(df_ranks$names, pattern = "nspecangle"), 
     unlist(lapply(strsplit(df_ranks$names, split = "_|[.]csv"), "[", 8)), df_ranks$tolerance)
df_ranks$tolerance <- ifelse(grepl(df_ranks$names, pattern = "cosine"), 
     unlist(lapply(strsplit(df_ranks$names, split = "_|[.]csv"), "[", 8)), df_ranks$tolerance)
df_ranks$tolerance <- ifelse(grepl(df_ranks$names, pattern = "navdis"), 
    unlist(lapply(strsplit(df_ranks$names, split = "_|[.]csv"), "[", 8)), df_ranks$tolerance)
df_ranks$tolerance <- ifelse(grepl(df_ranks$names, pattern = "neuclidean"), 
    unlist(lapply(strsplit(df_ranks$names, split = "_|[.]csv"), "[", 8)), df_ranks$tolerance)
df_ranks$tolerance <- ifelse(grepl(df_ranks$names, pattern = "Neutralloss"), 
    unlist(lapply(strsplit(df_ranks$names, split = "_|[.]csv"), "[", 2)), df_ranks$tolerance)
df_ranks$tolerance <- ifelse(grepl(df_ranks$names, pattern = "Spec2Vec"), NA, df_ranks$tolerance)
df_ranks$tolerance <- tolerance |>
    as.numeric()


## fill the algorithm
df_ranks$algorithm <- ifelse(grepl(df_ranks$names, pattern = "nspecangle"), 
    "Spectral_angle", df_ranks$algorithm)
df_ranks$algorithm <- ifelse(grepl(df_ranks$names, pattern = "cosine"), 
    "Cosine", df_ranks$algorithm)
df_ranks$algorithm <- ifelse(grepl(df_ranks$names, pattern = "navdis"), 
    "Absolute_distance", df_ranks$algorithm)
df_ranks$algorithm <- ifelse(grepl(df_ranks$names, pattern = "neuclidean"), 
    "Euclidean", df_ranks$algorithm)
df_ranks$algorithm <- ifelse(grepl(df_ranks$names, pattern = "Neutralloss"), 
    "Neutral_loss", df_ranks$algorithm)
df_ranks$algorithm <- ifelse(grepl(df_ranks$names, pattern = "Spec2Vec"), 
    "Spec2Vec", df_ranks$algorithm)


## what does the split explain in the distribution?
tol_0005 <- df_ranks |>
    filter(algorithm == "Spectral_angle" & tolerance == 0.005)
tol_001 <- df_ranks |>
    filter(algorithm == "Spectral_angle" & tolerance == 0.01)
boxplot(tol_0005$corr, tol_001$corr, names = c("0.005", "0.01"))

rmarkdown::paged_table(df_ranks)

## do the actual histogram plotting
df_ranks |>
    ggplot() +
    geom_histogram(aes(x = ranks, col = algorithm, fill = algorithm), 
        alpha = 0.2, position = "identity")
df_ranks |> 
    ggplot() +
    geom_density(aes(x = corr, col = algorithm, fill = algorithm), 
        alpha = 0.2, position = "identity")


## Scatterplot overlaying the top 5 algorithms
df_l |>
    filter(name %in% names(corr_spearman_tanimoto_sort[1:5])) |>
    ggplot() +
    geom_point(aes(x = tanimoto_scores, y = value, 
        col = name, group = name), alpha = 0.2) +
    theme_classic()
```


## Only matching

```{r only_matching}
## filter the matching
df <- df_all[df_all$matching, ]

table(df$tanimoto_scores)

## remove the pccid and the matching column
df <- df |>
    select(-c(pccid, matching))
```


## Correlation between similarity coefficients and Tanimoto

`cols` specifies the columns that contain the Spectral similarity scores, 
not the Tanimoto coefficients.

### Calculate correlation

```{r}
## for Pearson correlation
.cols_remove <- c("id_query", "ref_id") 
corr_pearson <- df |>
    select(-all_of(.cols_remove)) |>
    as.matrix() |>
    cor(method = "pearson", use = "complete.obs")

## for Spearman correlation
corr_spearman <- df |> 
    select(-all_of(.cols_remove)) |>
    as.matrix() |>
    cor(method = "spearman", use = "complete.obs")

## extract the information on the Tanimoto coefficient and delete the 
## respective columns and rows

.col <- "tanimoto_scores" ############################################
corr_pearson_tanimoto <- corr_pearson[, .col]
corr_spearman_tanimoto <- corr_spearman[, .col]

.ind_remove <- !colnames(corr_pearson) %in% .col
corr_pearson <- corr_pearson[.ind_remove, .ind_remove]
corr_spearman <- corr_spearman[.ind_remove, .ind_remove]
```


Add annotations for rows.

```{r}
.rn <- rownames(corr_pearson)
algorithm <- character(length = nrow(corr_pearson))
tolerance <- numeric(length = nrow(corr_pearson))
mz_power <- numeric(length = nrow(corr_pearson))
intensity_power <- numeric(length = nrow(corr_pearson))
allowed_missing_percentage <- rep(NA, length = nrow(corr_pearson))

## algorithm
algorithm <- ifelse(grepl("nspecangle", .rn), "Spectral_angle", algorithm)
algorithm <- ifelse(grepl("cosine", .rn), "Cosine", algorithm)
algorithm <- ifelse(grepl("navdis", .rn), "Absolute_distance", algorithm)
algorithm <- ifelse(grepl("neuclidean", .rn), "Euclidean", algorithm)
algorithm <- ifelse(grepl("Neutralloss", .rn), "Neutral_loss", algorithm)
algorithm <- ifelse(grepl("Spec2Vec", .rn), "Spec2Vec", algorithm)
algorithm <- algorithm |>
    as.character()

## tolerance
tolerance <- ifelse(grepl("nspecangle", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 8)), tolerance)
tolerance <- ifelse(grepl("cosine", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 8)), tolerance)
tolerance <- ifelse(grepl("navdis", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 8)), tolerance)
tolerance <- ifelse(grepl("neuclidean", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 8)), tolerance)
tolerance <- ifelse(grepl("Neutralloss", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 2)), tolerance)
tolerance <- ifelse(grepl("Spec2Vec", .rn), NA, tolerance)
tolerance <- tolerance |>
    as.numeric()

## mz_power
mz_power <- ifelse(grepl("nspecangle", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), mz_power)
mz_power <- ifelse(grepl("cosine", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), mz_power)
mz_power <- ifelse(grepl("navdis", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), mz_power)
mz_power <- ifelse(grepl("neuclidean", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), mz_power)
mz_power <- ifelse(grepl("Neutralloss", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 3)), mz_power)
mz_power <- ifelse(grepl("Spec2Vec", .rn), NA, mz_power)
mz_power <- mz_power |>
    as.numeric()

## intensity_power
intensity_power <- ifelse(grepl("nspecangle", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 6)), intensity_power)
intensity_power <- ifelse(grepl("cosine", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 6)), intensity_power)
intensity_power <- ifelse(grepl("navdis", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 6)), intensity_power)
intensity_power <- ifelse(grepl("neuclidean", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 6)), intensity_power)
intensity_power <- ifelse(grepl("Neutralloss", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), intensity_power)
intensity_power <- ifelse(grepl("Spec2Vec", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 2)), intensity_power)
intensity_power <- intensity_power |>
    as.numeric()


## allowed_missing_percentage
## Spec2Vec


column_ha <- HeatmapAnnotation(algorithm = algorithm)
row_ha <- rowAnnotation( 
    tolerance = as.factor(tolerance), 
    mz_power = as.factor(mz_power), 
    intensity_power = as.factor(intensity_power))

## Pearson coefficients
Heatmap(mat = corr_pearson, top_annotation = column_ha,
    right_annotation = row_ha, show_column_names = FALSE, show_row_names = FALSE)

## Spearman coefficients
Heatmap(mat = corr_spearman, top_annotation = column_ha,
    right_annotation = row_ha, show_column_names = FALSE, show_row_names = FALSE)
```


## Only non-matching

```{r}
## only keep the non-matching
df <- df_all[!df_all$matching, ]

hist(df$tanimoto_scores, main="Tanimoto coefficients (non-matching)")

## remove the pccid and the matching column
df <- df |>
    select(-c(pccid, matching))

```


## Correlation between similarity coefficients and Tanimoto

`cols` specifies the columns that contain the Spectral similarity scores, 
not the Tanimoto coefficients.

### Calculate correlation

```{r}
## for Pearson correlation
.cols_remove <- c("id_query", "ref_id") 
corr_pearson <- df |>
    select(-all_of(.cols_remove)) |>
    as.matrix() |>
    cor(method = "pearson", use = "complete.obs")

## for Spearman correlation
corr_spearman <- df |> 
    select(-all_of(.cols_remove)) |>
    as.matrix() |>
    cor(method = "spearman", use = "complete.obs")

## extract the information on the Tanimoto coefficient and delete the 
## respective columns and rows

.col <- "tanimoto_scores" ############################################
corr_pearson_tanimoto <- corr_pearson[, .col]
corr_spearman_tanimoto <- corr_spearman[, .col]

.ind_remove <- !colnames(corr_pearson) %in% .col
corr_pearson <- corr_pearson[.ind_remove, .ind_remove]
corr_spearman <- corr_spearman[.ind_remove, .ind_remove]
```


Add annotations for rows.

```{r}
.rn <- rownames(corr_pearson)
algorithm <- character(length = nrow(corr_pearson))
tolerance <- numeric(length = nrow(corr_pearson))
mz_power <- numeric(length = nrow(corr_pearson))
intensity_power <- numeric(length = nrow(corr_pearson))
allowed_missing_percentage <- rep(NA, length = nrow(corr_pearson))

## algorithm
algorithm <- ifelse(grepl("nspecangle", .rn), "Spectral_angle", algorithm)
algorithm <- ifelse(grepl("cosine", .rn), "Cosine", algorithm)
algorithm <- ifelse(grepl("navdis", .rn), "Absolute_distance", algorithm)
algorithm <- ifelse(grepl("neuclidean", .rn), "Euclidean", algorithm)
algorithm <- ifelse(grepl("Neutralloss", .rn), "Neutral_loss", algorithm)
algorithm <- ifelse(grepl("Spec2Vec", .rn), "Spec2Vec", algorithm)
algorithm <- algorithm |>
    as.character()

## tolerance
tolerance <- ifelse(grepl("nspecangle", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 8)), tolerance)
tolerance <- ifelse(grepl("cosine", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 8)), tolerance)
tolerance <- ifelse(grepl("navdis", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 8)), tolerance)
tolerance <- ifelse(grepl("neuclidean", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 8)), tolerance)
tolerance <- ifelse(grepl("Neutralloss", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 2)), tolerance)
tolerance <- ifelse(grepl("Spec2Vec", .rn), NA, tolerance)
tolerance <- tolerance |>
    as.numeric()

## mz_power
mz_power <- ifelse(grepl("nspecangle", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), mz_power)
mz_power <- ifelse(grepl("cosine", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), mz_power)
mz_power <- ifelse(grepl("navdis", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), mz_power)
mz_power <- ifelse(grepl("neuclidean", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), mz_power)
mz_power <- ifelse(grepl("Neutralloss", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 3)), mz_power)
mz_power <- ifelse(grepl("Spec2Vec", .rn), NA, mz_power)
mz_power <- mz_power |>
    as.numeric()

## intensity_power
intensity_power <- ifelse(grepl("nspecangle", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 6)), intensity_power)
intensity_power <- ifelse(grepl("cosine", .rn), 
     unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 6)), intensity_power)
intensity_power <- ifelse(grepl("navdis", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 6)), intensity_power)
intensity_power <- ifelse(grepl("neuclidean", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 6)), intensity_power)
intensity_power <- ifelse(grepl("Neutralloss", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 4)), intensity_power)
intensity_power <- ifelse(grepl("Spec2Vec", .rn), 
    unlist(lapply(strsplit(.rn, split = "_|[.]csv"), "[", 2)), intensity_power)
intensity_power <- intensity_power |>
    as.numeric()


## allowed_missing_percentage
## Spec2Vec


column_ha <- HeatmapAnnotation(algorithm = algorithm)
row_ha <- rowAnnotation( 
    tolerance = as.factor(tolerance), 
    mz_power = as.factor(mz_power), 
    intensity_power = as.factor(intensity_power))

## Pearson coefficients
Heatmap(mat = corr_pearson, top_annotation = column_ha,
    right_annotation = row_ha, show_column_names = FALSE, show_row_names = FALSE)

## Spearman coefficients
Heatmap(mat = corr_spearman, top_annotation = column_ha,
    right_annotation = row_ha, show_column_names = FALSE, show_row_names = FALSE)
```

### Scatterplots



Using `ggplot2`

```{r}
##sample_prop <- sample(1:nrow(df), ceiling(nrow(df) * 0.2), 
##    replace = FALSE)


df |>##[sample_prop, ] |>
    ggplot(aes_string(x = colnames(df)[3], y = colnames(df)[4])) +
    geom_point(cex = 0.2, alpha = 0.2) +
    geom_smooth(method = "lm") +
    theme_classic()

df |>##[sample_prop, ] |>
    ggplot(aes_string(x = colnames(df)[3], y = colnames(df)[99])) +
    geom_point(cex = 0.2, alpha = 0.2) +
    geom_smooth(method = "lm") +
    theme_classic()
```


### Tanimoto against spectral similarities

The last column (`ncol(df)`) contains the Tanimoto score.

```{r}
df_l <- pivot_longer(df, cols = 4:ncol(df))
```

Do the actual plotting. Only show the data for the top 5 scores

```{r}
corr_spearman_tanimoto_sort <- sort(corr_spearman_tanimoto, decreasing = TRUE)

corr_spearman_tanimoto_sort[1:10]


## remove the .col (Tanimoto column) with correlation of 1
corr_spearman_tanimoto_sort <- corr_spearman_tanimoto_sort[!names(corr_spearman_tanimoto_sort) %in% .col]

## create histograms per algorithm that show the rank of the correlation
df_ranks <- data.frame(
    corr = sort(corr_spearman_tanimoto_sort, decreasing = T),
    ranks = order(corr_spearman_tanimoto_sort, decreasing=T),
    names = names(corr_spearman_tanimoto_sort[order(corr_spearman_tanimoto_sort, decreasing=T)]),
    tolerance = NA,
    algorithm = "")

## fill the tolerance
df_ranks$tolerance <- ifelse(grepl(df_ranks$names, pattern = "nspecangle"), 
     unlist(lapply(strsplit(df_ranks$names, split = "_|[.]csv"), "[", 8)), df_ranks$tolerance)
df_ranks$tolerance <- ifelse(grepl(df_ranks$names, pattern = "cosine"), 
     unlist(lapply(strsplit(df_ranks$names, split = "_|[.]csv"), "[", 8)), df_ranks$tolerance)
df_ranks$tolerance <- ifelse(grepl(df_ranks$names, pattern = "navdis"), 
    unlist(lapply(strsplit(df_ranks$names, split = "_|[.]csv"), "[", 8)), df_ranks$tolerance)
df_ranks$tolerance <- ifelse(grepl(df_ranks$names, pattern = "neuclidean"), 
    unlist(lapply(strsplit(df_ranks$names, split = "_|[.]csv"), "[", 8)), df_ranks$tolerance)
df_ranks$tolerance <- ifelse(grepl(df_ranks$names, pattern = "Neutralloss"), 
    unlist(lapply(strsplit(df_ranks$names, split = "_|[.]csv"), "[", 2)), df_ranks$tolerance)
df_ranks$tolerance <- ifelse(grepl(df_ranks$names, pattern = "Spec2Vec"), NA, df_ranks$tolerance)
df_ranks$tolerance <- tolerance |>
    as.numeric()


## fill the algorithm
df_ranks$algorithm <- ifelse(grepl(df_ranks$names, pattern = "nspecangle"), 
    "Spectral_angle", df_ranks$algorithm)
df_ranks$algorithm <- ifelse(grepl(df_ranks$names, pattern = "cosine"), 
    "Cosine", df_ranks$algorithm)
df_ranks$algorithm <- ifelse(grepl(df_ranks$names, pattern = "navdis"), 
    "Absolute_distance", df_ranks$algorithm)
df_ranks$algorithm <- ifelse(grepl(df_ranks$names, pattern = "neuclidean"), 
    "Euclidean", df_ranks$algorithm)
df_ranks$algorithm <- ifelse(grepl(df_ranks$names, pattern = "Neutralloss"), 
    "Neutral_loss", df_ranks$algorithm)
df_ranks$algorithm <- ifelse(grepl(df_ranks$names, pattern = "Spec2Vec"), 
    "Spec2Vec", df_ranks$algorithm)


## what does the split explain in the distribution?
tol_0005 <- df_ranks |>
    filter(algorithm == "Spectral_angle" & tolerance == 0.005)
tol_001 <- df_ranks |>
    filter(algorithm == "Spectral_angle" & tolerance == 0.01)
boxplot(tol_0005$corr, tol_001$corr, names = c("0.005", "0.01"))

rmarkdown::paged_table(df_ranks)

## do the actual histogram plotting
df_ranks |>
    ggplot() +
    geom_histogram(aes(x = ranks, col = algorithm, fill = algorithm), 
        alpha = 0.2, position = "identity")
df_ranks |> 
    ggplot() +
    geom_density(aes(x = corr, col = algorithm, fill = algorithm), 
        alpha = 0.2, position = "identity")


## Scatterplot overlaying the top 5 algorithms
df_l |>
    filter(name %in% names(corr_spearman_tanimoto_sort[1:5])) |>
    ggplot() +
    geom_point(aes(x = tanimoto_scores, y = value, 
        col = name, group = name), alpha = 0.2) +
    theme_classic()
```







